Left-Linear Rewriting in Adhesive Categories

### 2024-07-09 concur/final.tex
sistemato spazio
aggiunto url arxiv
pass su concur

\documentclass[a4paper,UKenglish,cleveref,pdftex,thm-restate,numberwithinsect]{lipics-v2021}
\newif\ifreport
\reporttrue 
%\reportfalse

@@

\ifreport
\nolinenumbers %uncomment to disable line numbering
\hideLIPIcs  %uncomment to remove references to LIPIcs series (logo, DOI, ...), e.g. when preparing a pre-final version to be uploaded to arXiv or another public repository
\else
\relatedversion{An extended version is available at \url{https://arxiv.org/abs/2007.04213}.} %optional, e.g. full version hosted on arXiv, HAL, or other respository/website
%\relatedversion{A full version of the paper is available at \url{...}.}
\fi

@@

One of the key pay-off of concurrency theory is the idea that the behaviour of 
a computational device can be modelled by abstracting its sequences of steps
%performed by the device 
via a suitable equivalence. Intuitively, the equivalence 
captures when 
%sequential 
steps are causally unrelated and thus could be executed in any 
order, and possibly in parallel. 
%
Two seminal contributions to this line of research have been
Mazurkiewicz's traces~\cite{Mazurkiewicz86} and Winskel's event structures~\cite{NPW:PNES}.
%
Concerning formalisms based on the rewriting paradigm,
% i.e.~where the behaviour is dictated by a set of rules that have to be instantiated with the system at hand,
concurrency often boils down to having a notion of 
independence between two consecutive steps.
%thus viewing a sequence of steps as a concurrent computation. 
Noteworthy examples are the
interchange law~\cite{Mes92} in term rewriting and permutation 
equivalence~\cite{JJL80} in $\lambda$-calculi.

@@

Just to get some basic intuition about left-linear rewriting systems, their capabilities and the problems that can arise, assume we want to model a situation in which a coffee comes with a bag of white sugar and a bag of brown sugar. The situation is represented by the graph $G_0$ in Fig.~\ref{fi:coffee-rewriting}: the loop $\mathit{c}$ is the coffee, the two loops $b$ and $w$  represent the white and brown sugar. The corresponding nodes are connected to the coffee node by an $s$-labelled edge to indicate that they can be possibly added to the coffee. The available rules are in Fig.~\ref{fi:coffee-rules}. Rules $\rho_w$ and $\rho_b$ model the addition of white and brown sugar, respectively, to the coffee.
%
Note that adding one kind of sugar, say $x$, to the coffee is realised
by deleting loop $x$ and merging the corresponding node with the
coffee node so that a self-loop $s$ is created.

Once some sugar has been added, the coffee is ready for drinking. This is represented by rule $\rho_r$, which checks for the presence of sugar, represented by the loop $s$, in the coffee, and brings it to the $r$ state, ready for drinking.

@@

\emph{Synopsis.}
% The paper has the following structure. 
In \S\ref{sec:ade} we recall the basic 
definitions of DPO rewriting for $\mathcal{M}$-adhesive categories. 
In \S\ref{sec:equi} we present the key novelties of our proposal,
introducing the standard notion of independence and 
a novel axiomatic notion of switchability, showing that the
former fails to imply the latter for left-linear rewriting systems.
We then define a subclass of left-linear rewriting systems
that we call well-switching rewriting systems,
and we prove that the classical results for linear rewriting systems
are recovered for such a class.
Finally, in \S\ref{conclusions} we outline directions for future work.
%
The appendices
% collect additional data,
recall basic results on adhesive categories (Appendix~\ref{app:ade}),
present \emph{strong enforcing} rewriting systems, a class of systems
where the local Church-Rosser Theorem holds (Appendix~\ref{app:fill}),
and give proofs of the results
(Appendix~\ref{omitted}).

@@

\subsection{$\mathcal{M}$-adhesivity}\label{subsec:ade}
The key property of $\mathcal{M}$-adhesive categories is the \emph{Van Kampen condition}~\cite{brown1997van,johnstone2007quasitoposes,lack2005adhesive}. % We
% will recall it and examine some of its consequences.
In order to define it we need to introduce some terminology.
% and facts regarding subclasses of
% $\mor(\X)$.
%
%\begin{definition}
  Let  $\X$ be a category. A subclass $\mathcal{A}$ of
  $\mor(\X)$ is called %\todo{il diagramma confonde la composizione}
  \begin{itemize}
    \parbox{11cm}{\item
      \emph{stable under pushouts (pullbacks)} if for every pushout (pullback) square as the one on the right, 	if $m \in \mathcal{A}$ ($n\in \mathcal{A}$) then $n \in \mathcal{A}$ ($m \in \mathcal{A}$);
    \item \emph{closed under composition} if $h, k\in \mathcal{A}$ implies $h\circ k\in \mathcal{A}$ whenever $h$ and $k$ are composable.{\tiny }}\hfill
    \parbox{2cm}{$\xymatrix{A\ar[r]^f  \ar[d]_{m}& B \ar[d]^n \\ C \ar[r]_g & D}$}
    \parbox{11cm}{}\hfill
  \end{itemize}
%\end{definition}

@@

%It can be seen that 
$\mathcal{M}$-adhesivity of $\X$ ensures the
uniqueness of the result of applying a rule to an object: two
direct derivations using the same match are abstraction equivalent
(see Proposition~\ref{prop:unique}).

@@

It can be shown that switches along the same independence pair are unique up to abstraction
equivalence (see Proposition~\ref{thm:switch_uni}).

@@

The fact that sequential independence implies switchability always
holds for linear rules (see \Cref{prop:equi}). The result is so
indispensable that typically, in the literature, it is not even
stated, in the sense that switchability is not introduced
axiomatically as in Definition~\ref{def:switch}, but it is based on
the explicit construction of a switch.

@@

\begin{definition}[strong independence pair]
  \label{def:filler}
  Let $\X$ be an $\mathcal{M}$-adhesive category 
  and $(\X, \R)$ a left-linear rewriting system. 
  Let also $(i_0, i_1)$ be an independence pair between two direct
  derivations $\dder{D}_0$, $\dder{D}_1$
  as in the solid part of the diagram below
  \[
    \xymatrix@C=18pt@R=16pt{L_0 \ar[d]_{m_0}&& K_0
      \ar[d]_{k_0}\ar@{>->}[ll]_{l_0} \ar[r]^{r_0} \ar@{.>}@/^.20cm/[ddrr]|(.32)\hole|(.57)\hole_(.7){u_0} & R_0
      \ar@/^.35cm/[drrr]|(.3)\hole^(.2){i_0} \ar[dr]|(.3)\hole_{h_0}
      && L_1 \ar@/_.35cm/[dlll]_(.2){i_1} \ar[dl]|(.3)\hole^{m_1}& K_1
      \ar[d]^{k_1}\ar@{>->}[l]_{l_1} \ar[rr]^{r_1} \ar@{.>}@/_.20cm/[ddll]|(.32)\hole|(.57)\hole^(.7){u_1} && R_1 \ar[d]^{h_1} \\
      %
      G_0 &&
      \ar@{>->}[ll]^{f_0} D_0 \ar[rr]_(.3){g_0}&& G_1 &&
      \ar@{>->}[ll]^(.3){f_1} D_1 \ar[rr]_{g_1}&& G_2\\
      %
      & & & & P \ar@/^.3cm/@{>.>}[ull]^{p_0} \ar@/_.3cm/@{.>}[urr]_{p_1}
    }
  \]

  Consider the pullback of $g_0$ and $f_1$, which yields $p_i : P \to D_i$
  for $i \in \{0,1\}$ and the mediating arrows $u_i\colon K_i\to P$ 
  for $i \in \{0,1\}$ 
  %and $u_1\colon K_1\to P$ 
  into the pullback object (see \Cref{prop:tec} for details). 
  % given by \Cref{prop:tec} .
  %
  We say that $(i_0, i_1)$ is a \emph{strong independence pair} if
  the first two squares depicted  
  below are pushouts and if the pushout of $r_1 : K_1 \to R_1$ and $u_1 : K_1 \to P$ exists
  \[
    \xymatrix@R=16pt{
      K_0 \ar[r]^{r_0} \ar[d]_{u_0}& R_0 \ar[d]^{i_0} & K_1
      \ar@{>->}[r]^{l_1} \ar[d]_{u_1}& L_1 \ar[d]^{i_1}
       &K_1 \ar[r]^{r_1} \ar[d]_{u_1}& R_1\ar@{.>}[d]^{j_0}
      \\
      %
      P \ar[r]_{p_1} & D_1 &P \ar[r]_{p_0} & D_0
       & P \ar@{.>}[r]_{q_1} & Q_1
    }
  \]
\end{definition}

@@

Consistency with the theory of linear rewriting systems is ensured by the fact that all linear rewriting systems are strong enforcing (see \cref{prop:equi}).

@@

% The previous definition might sound restrictive.
Clearly, %It is immediate to see that 
linear rewriting systems are well-switching (see Proposition~\ref{pr:weak}).
%
Moreover, we next observe that various classes of rewriting systems,
comprising all the ones used in modelling the applications to the
encoding of process calculi or of bio and chemical systems mentioned
in the introduction, are actually well-switching.

@@

When dealing with left-linear rules, such a characterisation
fails. Leaving aside the formal details (the interested reader can
refer to \Cref{def:permcon}), the intuition is quite simple. In the
linear case one proves that two derivations using the same rules in
different orders are switch equivalent when the colimits of the two
derivations, seen as diagrams, are isomorphic and the isomorphism
properly commutes with the embeddings of the rules. When considering
left-linear derivations, this is no longer true. Intuitively this is
due to the fact that a rewriting step can merge parts of an
object thus making the colimit little informative.
%
This effect can be seen for derivations $\der{D}$ in
Fig.~\ref{fi:derD} and $\der{D}'$ in Fig.~\ref{fi:derD1}. They are not
abstraction equivalent but the colimit of both derivations is a single node with two self-loops, in a way that  all commutativity requirements are
trivially satisfied.  
% OLD VERSION
% This is readily seen by considering derivations $\der{D}$
% in Fig.~\ref{fi:derD} and $\der{D}'$ in Fig.~\ref{fi:derD1}. We
% observed that $\der{D}$ and $\der{D}'$ are not switch equivalent.
% % 
% Still, it is easy to see that they are related by a consistent\todo{le definiamo solo nell'appendice}
% permutation, which is the identity on productions. The colimit of
% both derivations is a single node with a self-loop, i.e.~the
% terminal object in $\cat{Graph}$, hence all commutativity requirements
% are trivially satisfied.

@@

\newpage
Still, we can identify a
canonical form for switching derivations: at each step one can apply a switch to the inversion with largest index.

@@

Working in an $\mathcal{M}$-adhesive category we can guarantee that pushout complements are unique.

@@

\begin{theorem}\label{thm:good1}Let $\X$ be an $\mathcal{M}$-adhesive category with all pushouts and satisfying the $\mathcal{M}$-mixed and the $\mathcal{M}$-pushout decomposition properties. Then every left-linear rewriting system on $\X$ is strong enforcing.
\end{theorem}
\begin{proof}
	Consider the derivation $\der{D}=\{\dder{D}_i\}_{i=0}^1$, made by two sequentially independent derivations, depicted below
	\[\xymatrix@C=15pt{L_0 \ar[d]_{m_0}&& K_0 \ar[d]_{k_0}\ar@{>->}[ll]_{l_0} \ar[r]^{r_0} & R_0 \ar@/^.35cm/[drrr]|(.3)\hole_(.4){i_0} \ar[dr]|(.3)\hole_{h_0} && L_1 \ar@/_.35cm/[dlll]^(.4){i_1} \ar[dl]|(.3)\hole^{m_1}& K_1 \ar[d]^{k_1}\ar@{>->}[l]_{l_1} \ar[rr]^{r_1} && R_1 \ar[d]^{h_1} \\G_0 && \ar@{>->}[ll]^{f_0} D_0 \ar[rr]_{g_0}&& G_1  && \ar@{>->}[ll]^{f_1} D_1 \ar[rr]_{g_1}&& G_2 }\]

@@

\begin{definition} 
	Let $\X$ be an  $\mathcal{M}$-adhesive category, the pair $(\X, \mathcal{M})$ is of \emph{type $\mathbb{B}$} if
	\begin{enumerate}
		\item every arrow in $\mathcal{M}$ is a coproduct coprojection;
		\item $\X$ has all pushouts;
		\item $\X$ has strict initial objects and, for every object $X$, the unique arrow $?_X\colon 0\to X $ belongs to $\mathcal{M}$;
		\item all pushouts are $\mathcal{M}$-stable.
	\end{enumerate}
\end{definition}

\begin{remark}
	It is worth to examine more closely conditions $1$ and $3$ of the above definition.
	\smallskip 
	\begin{itemize}
	\parbox{10.3cm}{\item Let $m_0\colon X_0 \rightarrowtail Y$ be an arrow in $\mathcal{M}$, the first condition means that there exists $m_1\colon X_1\to Y$ such that $(Y, \{m_i\}_{i=0}^1)$ is a coproduct. This, together with property $3$, entails that every coprojection in a coproduct is in $\mathcal{M}$. This follows since any coproduct $(X_1+X_2, \{\iota_{X_i}\}_{i=0}^1)$ fits in a pushout diagram as the one aside.}
	  \parbox{3cm}{\vspace{-0em}$\xymatrix@C=25pt{0  \ar@{>->}[r]^-{?_{X_1}} \ar@{>->}[d]^{?_{X_2}}& X_1 \ar@{>->}[d]^{\iota_{X_1}}\\ X_2 \ar@{>->}[r]_-{\iota_{X_2}} & X_1+X_2}$}

@@

\begin{example}
The category $\Set$ of sets and functions, with its class of monos, is of type $\mathbb{B}$. Similarly, the category $\textbf{Inj}$ of sets and injective functions is quasiadhesive and, with its class of regular monos, of type $\mathbb{B}$. 
\end{example}

Categories of type $\mathbb{B}$ satisfy a property resembling \emph{extensivity} \cite{carboni1993introduction}.

@@

\begin{definition} Let $\X$ be an $\mathcal{M}$-adhesive category, we say that the pair $(\X, \mathcal{M})$ is of \emph{type $\mathbb{B}^+$} if (at least) one of the following holds
	\begin{enumerate}
		\item $(\X, \mathcal{M})$ is of type $\mathbb{B}$;
	\item $\X$ is $\Y/Y$ and $\mathcal{M}=\mathcal{N}/Y$for some $(\Y, \mathcal{N})$ of type $\mathbb{B}^+$ and $Y\in \Y$;
 	\item $\X$ is $Y/\Y$ and $\mathcal{M}=Y/\mathcal{N}$for some $(\Y, \mathcal{N})$ of type $\mathbb{B}^+$ and $Y\in \Y$:
 	\item $\X$ is $\Y^{\A}$ and $\mathcal{M}=\mathcal{N}^{\A}$ for some category category $\A$ and $(\Y, \mathcal{N})$ of type $\mathbb{B}^+$.
	\end{enumerate}
\end{definition}
\begin{example}
  \label{ex:graph-type-bb}
	The pair $(\mathbf{Graph}, \mon(\mathbf{Graph}))$ made by the topos of graphs and its class of monos is of type $\mathbb{B}^+$  but not of type $\mathbb{B}$. Indeed, not every monomorphism of graphs is a coproduct coprojection.
\end{example}

>>>>

\documentclass[a4paper,UKenglish,cleveref,pdftex,thm-restate,numberwithinsect]{lipics-v2021}
\newif\ifreport
%\reporttrue 
\reportfalse

@@

\ifreport
\nolinenumbers %uncomment to disable line numbering
\hideLIPIcs  %uncomment to remove references to LIPIcs series (logo, DOI, ...), e.g. when preparing a pre-final version to be uploaded to arXiv or another public repository
\else
\relatedversion{An extended version is available at \url{https://arxiv.org/abs/2407.06181}.} %optional, e.g. full version hosted on arXiv, HAL, or other respository/website
%\relatedversion{A full version of the paper is available at \url{...}.}
\fi

@@

One of the key pay-off of concurrency theory is the idea that the behaviour of 
a computational device can be modelled by abstracting its sequences of steps
%performed by the device 
via a suitable equivalence. Intuitively, the equivalence 
captures when 
%sequential 
steps are causally unrelated and thus could be executed in any 
order, and possibly in parallel. 
%
Two seminal contributions to this line of research have been
Mazurkiewicz's traces~\cite{Mazurkiewicz86} and Winskel's event structures~\cite{NPW:PNES}.
%
Concerning formalisms based on the rewriting paradigm,
% i.e.~where the behaviour is dictated by a set of rules that have to be instantiated with the system at hand,
concurrency often boils down to having a notion of 
independence between two consecutive steps.
%thus viewing a sequence of steps as a concurrent computation. 
Noteworthy examples are the
interchange law~\cite{Mes92} in term rewriting and permutation 
equivalence~\cite{JJL80} in $\lambda$-calculi. 
%\newpage

@@

Just to get some basic intuition about left-linear rewriting systems, their capabilities and the problems that can arise, assume we want to model a situation in which a coffee comes with a bag of white sugar and a bag of brown sugar. The situation is represented by the graph $G_0$ in Fig.~\ref{fi:coffee-rewriting}: the loop $\mathit{c}$ is the coffee, the two loops $b$ and $w$  represent the white and brown sugar. The corresponding nodes are connected to the coffee node by an $s$-labelled edge to indicate that they can be possibly added to the coffee. The available rules are in Fig.~\ref{fi:coffee-rules}. Rules $\rho_w$ and $\rho_b$ model the addition of white and brown sugar, respectively, to the coffee.
%
Note that adding one kind of sugar, say $x$, to the coffee is realised
by deleting loop $x$ and merging the corresponding node with the
coffee node so that a self-loop $s$ is created.
%
Once some sugar has been added, the coffee is ready for drinking. 
This is represented by rule $\rho_r$, which checks for the presence of sugar, represented by the loop $s$, in the coffee, 
and brings it to the $r$ state, ready for drinking.

@@

\emph{Synopsis.}
% The paper has the following structure. 
In \S\ref{sec:ade} we recall the basic 
definitions of DPO rewriting for $\mathcal{M}$-adhesive categories. 
In \S\ref{sec:equi} we present the key novelties of our proposal,
introducing the standard notion of independence and 
a novel axiomatic notion of switchability, showing that the
former fails to imply the latter for left-linear rewriting systems.
We then define well-switching rewriting systems,
a subclass of left-linear rewriting systems,
and we prove that the classical results for linear rewriting systems
are recovered for such a class.
Finally, in \S\ref{conclusions} we outline directions for future work.
%
%The appendices
% collect additional data,
Appendix~\ref{app:ade} recalls basic results on adhesive categories and
Appendix~\ref{app:fill}
presents \emph{strong enforcing} rewriting systems, a class of systems
where the local Church-Rosser Theorem holds.
Appendix C of the
extended version provides the proofs
of our results.

@@

\subsection{$\mathcal{M}$-adhesivity}\label{subsec:ade}
The key property of $\mathcal{M}$-adhesive categories is the \emph{Van Kampen condition}~\cite{brown1997van,johnstone2007quasitoposes,lack2005adhesive}. % We
% will recall it and examine some of its consequences.
%In order to define it we need to introduce some terminology.
% and facts regarding subclasses of
% $\mor(\X)$.
%
%\begin{definition}
  Let  $\X$ be a category. A subclass $\mathcal{A}$ of
  $\mor(\X)$ is called %\todo{il diagramma confonde la composizione}
  \begin{itemize}
    \parbox{11cm}{\item
      \emph{stable under pushouts (pullbacks)} if for every pushout (pullback) square as the one on the right, 	if $m \in \mathcal{A}$ ($n\in \mathcal{A}$) then $n \in \mathcal{A}$ ($m \in \mathcal{A}$);
    \item \emph{closed under composition} if $h, k\in \mathcal{A}$ implies $h\circ k\in \mathcal{A}$ whenever $h$ and $k$ are composable.{\tiny }}\hfill
    \parbox{2cm}{$\xymatrix{A\ar[r]^f  \ar[d]_{m}& B \ar[d]^n \\ C \ar[r]_g & D}$}
    \parbox{11cm}{}\hfill
  \end{itemize}
%\end{definition}

@@

%It can be seen that 
$\mathcal{M}$-adhesivity ensures the
uniqueness of the result of applying a rule to an object: two
direct derivations using the same match are abstraction equivalent
(see Proposition~C.1 of the extended version).

@@

It can be shown that switches along the same independence pair are unique up to abstraction
equivalence (see Proposition~C.3 of the extended version).

@@

The fact that sequential independence implies switchability always
holds for linear rules (see Proposition~C.8 of the extended version). The result is so
indispensable that typically, in the literature, it is not even
stated, in the sense that switchability is not introduced
axiomatically as in Definition~\ref{def:switch}, but it is based on
the explicit construction of a switch.

@@

\begin{definition}[strong independence pair]
  \label{def:filler}
  Let $\X$ be an $\mathcal{M}$-adhesive category 
  and $(\X, \R)$ a left-linear rewriting system. 
  Let also $(i_0, i_1)$ be an independence pair between two direct
  derivations $\dder{D}_0$, $\dder{D}_1$
  as in the solid part of the diagram below
  \[
    \xymatrix@C=18pt@R=17pt{L_0 \ar[d]_{m_0}&& K_0
      \ar[d]_{k_0}\ar@{>->}[ll]_{l_0} \ar[r]^{r_0} \ar@{.>}@/^.20cm/[ddrr]|(.32)\hole|(.57)\hole_(.7){u_0} & R_0
      \ar@/^.35cm/[drrr]|(.3)\hole^(.2){i_0} \ar[dr]|(.3)\hole_{h_0}
      && L_1 \ar@/_.35cm/[dlll]_(.2){i_1} \ar[dl]|(.3)\hole^{m_1}& K_1
      \ar[d]^{k_1}\ar@{>->}[l]_{l_1} \ar[rr]^{r_1} \ar@{.>}@/_.20cm/[ddll]|(.32)\hole|(.57)\hole^(.7){u_1} && R_1 \ar[d]^{h_1} \\
      %
      G_0 &&
      \ar@{>->}[ll]^{f_0} D_0 \ar[rr]_(.3){g_0}&& G_1 &&
      \ar@{>->}[ll]^(.3){f_1} D_1 \ar[rr]_{g_1}&& G_2\\
      %
      & & & & P \ar@/^.3cm/@{>.>}[ull]^{p_0} \ar@/_.3cm/@{.>}[urr]_{p_1}
    }
  \]

  Consider the pullback of $g_0$ and $f_1$, which yields $p_i : P \to D_i$
  for $i \in \{0,1\}$ and the mediating arrows $u_i\colon K_i\to P$ 
  for $i \in \{0,1\}$ 
  %and $u_1\colon K_1\to P$ 
  into the pullback object (see Proposition~C.4 of the extended version for details). 
  % given by \Cref{prop:tec} .
  %
  We say that $(i_0, i_1)$ is a \emph{strong independence pair} if
  the first two squares depicted  
  below are pushouts and if the pushout of $r_1 : K_1 \to R_1$ and $u_1 : K_1 \to P$ exists
  \[
    \xymatrix@R=16pt{
      K_0 \ar[r]^{r_0} \ar[d]_{u_0}& R_0 \ar[d]^{i_0} & K_1
      \ar@{>->}[r]^{l_1} \ar[d]_{u_1}& L_1 \ar[d]^{i_1}
       &K_1 \ar[r]^{r_1} \ar[d]_{u_1}& R_1\ar@{.>}[d]^{j_0}
      \\
      %
      P \ar[r]_{p_1} & D_1 &P \ar[r]_{p_0} & D_0
       & P \ar@{.>}[r]_{q_1} & Q_1
    }
  \]
\end{definition}

@@

Consistency with the theory of linear rewriting systems is ensured by the fact that all linear rewriting systems are strong enforcing (see Proposition~C.8 of the extended version).

@@

% The previous definition might sound restrictive.
Clearly, %It is immediate to see that 
linear rewriting systems are well-switching (see Proposition~C.10 of the extended version).
%
Moreover, we next observe that various classes of rewriting systems,
comprising all the ones used in modelling the applications to the
encoding of process calculi or of bio and chemical systems mentioned
in the introduction, are actually well-switching.

@@

When dealing with left-linear rules, such a characterisation
fails. Leaving aside the formal details (the interested reader can
refer to Definition~C.12 of the extended version), the intuition is quite simple. In the
linear case one proves that two derivations using the same rules in
different orders are switch equivalent when the colimits of the two
derivations, seen as diagrams, are isomorphic and the isomorphism
properly commutes with the embeddings of the rules. When considering
left-linear derivations, this is no longer true. Intuitively this is
due to the fact that a rewriting step can merge parts of an
object thus making the colimit little informative.
%
This effect can be seen for derivations $\der{D}$ in
Fig.~\ref{fi:derD} and $\der{D}'$ in Fig.~\ref{fi:derD1}. They are not
abstraction equivalent but the colimit of both derivations is a single node with two self-loops, in a way that  all commutativity requirements are
trivially satisfied.  
% OLD VERSION
% This is readily seen by considering derivations $\der{D}$
% in Fig.~\ref{fi:derD} and $\der{D}'$ in Fig.~\ref{fi:derD1}. We
% observed that $\der{D}$ and $\der{D}'$ are not switch equivalent.
% % 
% Still, it is easy to see that they are related by a consistent\todo{le definiamo solo nell'appendice}
% permutation, which is the identity on productions. The colimit of
% both derivations is a single node with a self-loop, i.e.~the
% terminal object in $\cat{Graph}$, hence all commutativity requirements
% are trivially satisfied.

@@

Still, we can identify a
canonical form for switching derivations: at each step one can apply a switch to the inversion with largest index.

@@

Working in an $\mathcal{M}$-adhesive category guarantees that pushout complements are unique.

@@

\begin{theorem}\label{thm:good1}Let $\X$ be an $\mathcal{M}$-adhesive category with all pushouts and satisfying the $\mathcal{M}$-mixed and the $\mathcal{M}$-pushout decomposition properties. Then every left-linear rewriting system on $\X$ is strong enforcing.
\end{theorem}
\begin{proof}
	Let $\der{D}=\{\dder{D}_i\}_{i=0}^1$ be the derivation made by two sequentially independent derivations
	\[\xymatrix@C=15pt{L_0 \ar[d]_{m_0}&& K_0 \ar[d]_{k_0}\ar@{>->}[ll]_{l_0} \ar[r]^{r_0} & R_0 \ar@/^.35cm/[drrr]|(.3)\hole_(.4){i_0} \ar[dr]|(.3)\hole_{h_0} && L_1 \ar@/_.35cm/[dlll]^(.4){i_1} \ar[dl]|(.3)\hole^{m_1}& K_1 \ar[d]^{k_1}\ar@{>->}[l]_{l_1} \ar[rr]^{r_1} && R_1 \ar[d]^{h_1} \\G_0 && \ar@{>->}[ll]^{f_0} D_0 \ar[rr]_{g_0}&& G_1  && \ar@{>->}[ll]^{f_1} D_1 \ar[rr]_{g_1}&& G_2 }\]

@@

\begin{definition} 
	Let $\X$ be an  $\mathcal{M}$-adhesive category, the pair $(\X, \mathcal{M})$ is of \emph{type $\mathbb{B}$} if
	\begin{enumerate}
		\item every arrow in $\mathcal{M}$ is a coproduct coprojection;
		\item $\X$ has all pushouts;
		\item $\X$ has strict initial objects and, for every $X$, the unique arrow $?_X\colon 0\to X $ belongs to $\mathcal{M}$;
		\item all pushouts are $\mathcal{M}$-stable.
	\end{enumerate}
\end{definition}

\begin{remark}
	It is worth to examine more closely conditions $1$ and $3$ of the above definition.

	\vspace{-.1cm}
	\begin{itemize}
	\parbox{10.3cm}{\item Let $m_0\colon X_0 \rightarrowtail Y$ be in $\mathcal{M}$, the first condition means that there exists $m_1\colon X_1\to Y$ such that $(Y, \{m_i\}_{i=0}^1)$ is a coproduct. This, together with property $3$, entails that every coprojection in a coproduct is in $\mathcal{M}$. This follows since any coproduct $(X_1+X_2, \{\iota_{X_i}\}_{i=0}^1)$ fits in a pushout diagram as the one aside.
	}
	  \parbox{3cm}{\vspace{-0em}$\xymatrix@C=25pt{0  \ar@{>->}[r]^-{?_{X_1}} \ar@{>->}[d]^{?_{X_2}}& X_1 \ar@{>->}[d]^{\iota_{X_1}}\\ X_2 \ar@{>->}[r]_-{\iota_{X_2}} & X_1+X_2}$}

@@

\begin{example}
The category $\Set$ of sets and functions
%with its class of monos, 
is of type $\mathbb{B}$. Similarly, the category $\textbf{Inj}$ of sets and injective functions is quasiadhesive and, with regular monos, of type $\mathbb{B}$. 
\end{example}

\newpage
Categories of type $\mathbb{B}$ satisfy a property resembling \emph{extensivity} \cite{carboni1993introduction}.

@@

\begin{definition} Let $\X$ be an $\mathcal{M}$-adhesive category, we say that the pair $(\X, \mathcal{M})$ is of \emph{type $\mathbb{B}^+$} if (at least) one of the following holds
	\begin{enumerate}
		\item $(\X, \mathcal{M})$ is of type $\mathbb{B}$;
	\item $\X$ is $\Y/Y$ and $\mathcal{M}=\mathcal{N}/Y$for some $(\Y, \mathcal{N})$ of type $\mathbb{B}^+$ and $Y\in \Y$;
 	\item $\X$ is $Y/\Y$ and $\mathcal{M}=Y/\mathcal{N}$for some $(\Y, \mathcal{N})$ of type $\mathbb{B}^+$ and $Y\in \Y$:
 	\item $\X$ is $\Y^{\A}$ and $\mathcal{M}=\mathcal{N}^{\A}$ for some category category $\A$ and $(\Y, \mathcal{N})$ of type $\mathbb{B}^+$.
	\end{enumerate}
\end{definition}
\begin{example}
  \label{ex:graph-type-bb}
	The pair $(\mathbf{Graph}, \mon(\mathbf{Graph}))$ of graphs and their monos is of type $\mathbb{B}^+$  but not of type $\mathbb{B}$. Indeed, not every monomorphism of graphs is a coproduct coprojection.
\end{example}
